\documentclass[12pt]{book}
\usepackage{geometry}
\geometry{letterpaper, left=22.5mm, right=22.5mm, top=30mm, bottom=30mm}
\geometry{letterpaper}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{enumitem}
\usepackage{fancyhdr}
\usepackage{framed}
\usepackage{tikz}
\usepackage{mathpazo}
%\usepackage{charter}
%\usepackage{newcent}
\usepackage{indentfirst}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{float}
\usepackage{makecell}
\usepackage{xcolor}
\usepackage{mdframed}
\usetikzlibrary{trees}
\pagestyle{fancy}
\usepackage{amsthm}
\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]
\theoremstyle{property}
\newtheorem{property}{Property}[section]
\theoremstyle{assumption}
\newtheorem{assumption}{Assumption}[section]
\theoremstyle{example}
\newtheorem{example}{Example}[section]
\theoremstyle{comment}
\newtheorem{comment}{Comment}[section]
\newtheorem{theorem}{Theorem}[section]
\newtheorem{corollary}{Corollary}[theorem]
\newtheorem{lemma}[theorem]{Lemma}
\usepackage{lastpage}
\usepackage{wrapfig}
\usepackage{hyperref}
\usepackage{subcaption}
\usepackage{setspace}
\usepackage{natbib}
\hypersetup{
colorlinks=true,
linkcolor=black,
filecolor=green, 
urlcolor=blue,
}
\newcommand{\ROM}[1]
    {\MakeUppercase{\romannumeral #1}}
\fancyhead[L]{Econometrics \ROM{2} }%change each reci
\fancyhead[R]{Spring 2020}
\fancyfoot[C]{\thepage \hspace{1pt} / \pageref{LastPage}}

    \newcommand{\prefacename}{Preface}
    \newenvironment{preface}{
        \vspace*{\stretch{2}}
        {\noindent \bfseries \Huge \prefacename}
        \begin{center}
            % \phantomsection \addcontentsline{toc}{chapter}{\prefacename} % enable this if you want to put the preface in the table of contents
            \thispagestyle{plain}
        \end{center}%
    }
    {\vspace*{\stretch{5}}}

\fancypagestyle{firstpage}{%
\fancyhf{}%
\renewcommand{\headrulewidth}{0mm}%
  \fancyfoot[C]{\thepage \hspace{1pt} / \pageref{LastPage}}
}
%change title each rec

\title{Introduction to Econometrics \ROM{2}: The Book\footnote{This book is meant to leave what was learned in the spring semester of the year 2020 as a record, hoping to help future first-year students navigate through the Econometrics sequence. I do not intend to publish this in any form. Thus, while sharing this file with others not necessarily affiliated to Columbia is allowed, I forbid reproducing this book for commercial purposes.}}
\date{Spring 2020 Semester}
\author{Seung-hun Lee\footnote{sl4436@columbia.edu. Contact me if you find any errors. }}


\begin{document}
\linespread{1.25}
\onehalfspacing
\maketitle
\pagestyle{empty}

\frontmatter
\pagenumbering{roman}

\begin{preface}
This `Book' is a compilation of all the recitation notes that I have used in the Spring 2020 semester when I worked as a teaching fellow for the Introduction to Econometrics \ROM{2} class at Columbia University. The intention behind The Book is to create a one-stop reference for all things that are related to graduate-level econometrics.  In particular, this book focuses on the following topics
\begin{enumerate}
\item Instrumental Variables
\item Generalized Method of Moments
\item Panel Data (both static and dynamic)
\item Bootstrapping and Multiple Tests
\item Quantile Regression
\item Nonparametric and Semiparametric Estimation
\item Program Evaluation and Treatment Effects: Random Assignment, Selection on Observables, Selection on Unobservables, and Regression Discontinuity
\item Introduction to Machine Learning
\end{enumerate}
\par
In creating this book, I have been indebted to many individuals. First, the contents are mainly based on the lecture notes of Professors Jushan Bai and Bernard Salanie. I was given a rare opportunity to go over the first-year material once more. With this opportunity, I was able to gain mastery of the relevant methods that I can take anywhere in my research career. I am grateful to both professors for giving me this chance. 
\par
I was also greatly helped by two individuals - Paul Sungwook Koh and Dong Woo Hahm. They provided me an access to their prevous recitation notes. I was able to refine my explanation based on their notes, which benefitted those who learned from me and myself. Without them, this teaching fellowship would have been monumenally difficult. 
\par
Last but not least, I thank the first year cohort of 2019-2020 academic year. Not only have they corrected mistakes in my notes as classes went on, but they also gave me constructive and positive feedbacks for the recitation notes that became this book. The motivation from these reviews helped me complete this work. 
\par
Any remaining errors - typos, concepts that could be explained more efficiently than stated, etc. - are mine. I would appreciate any comments regarding the errors you may find here. \par

\end{preface}

\tableofcontents



\mainmatter
\pagestyle{fancy}
\input{Review.tex}
\input{IV.tex}
\input{GMM.tex}
\input{Panel.tex}
\input{BootMult.tex}
\input{Quantiles.tex}
\input{Nonparametrics.tex}
\input{TE1.tex}
\input{TE2.tex}
\input{TE3.tex}
\input{ML.tex}
\end{document}
